---
title: PRINCIPAL COMPONENT ANALYSIS ON USCRIMES DATA
output: word_document
Authors: Beenapreet
---

```{r setup, include=FALSE}

#knitr::opts_chunk$set(echo = TRUE)
```


```{r 1}
#setwd("C:/Users/beena/Downloads/Analytics Modelling")

uscrime<-read.table("uscrime.txt",header=TRUE)
View(uscrime)
```

Applying PCA to US Crimes dataset
```{r 2}
uscrime_pca<- prcomp(uscrime[1:15],scale = TRUE)
#### Viewing the First 4 principal components
head(uscrime_pca$x[,1:4])
summary(uscrime_pca)
```

```{r 3}
plot(uscrime_pca,type="line")
```

From the summary  and plot we can see that the first 4 principle componensts explain most of the variability of the data, hence we use the first 4 components to create our model
```{r 4}
### Extracting the first 4 principle components
PCA_uscrimes<-as.data.frame(cbind(uscrime_pca$x[,1],uscrime_pca$x[,2],uscrime_pca$x[,3],uscrime_pca$x[,4],uscrime$Crime))
#### linear regression model with principal components
model_pca<-lm(V5~.,data=PCA_uscrimes )
summary(model_pca)

```
Now we calculate the coefficients in terms of the original variables in our model using the eigen vectors from PCA. We iterate across the top 4 PCA coefficients and multiply the beta coefficients obtained from the linear regression model with each of the top 4 PCA values. 
```{r 5}
transformed_coeff<-c()
for(x in 1:(ncol(uscrime)-1)) {
  iter <- 0
  
  
  for (i in 1:4) {
    
    iter <- iter + model_pca$coefficients[i+1]*uscrime_pca$rotation[x,i]
    
  }
  
  
  transformed_coeff <- rbind(transformed_coeff, c(colnames(uscrime)[x],iter))
}
transformed_coeff <- as.data.frame(transformed_coeff)
colnames(transformed_coeff) <- c("Variable", "PCA Coefficient")
print(as.data.frame(transformed_coeff))
```


Explanation:

We get a R^2 value of 30.91% from the multiple regression model using 4 Principle components whereas from the previous question we get R^2 value of 76.59%. We see that the PCA model with 4 components does not explain as much of the variability hence R^2 is low as compared to the Multiple Linear regression model of the previous question.
